Understanding Recovery from Malnutrition in Early Life:\\ A Multi-Omic Analysis of the Infant Gut Microbiome, Plasma Lipidome, and Neurodevelopmental Outcomes

### 2025-08-14 writing/manuscript/main.tex
Manuscript updates
Updated figure 1

\documentclass{article}
\usepackage{float}
\usepackage[a4paper, top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage[super,sort&compress]{natbib}
\bibliographystyle{unsrtnat}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{booktabs} % Recommended for cleaner, professional-looking tables

@@

\newpage
\section{Main}
% Importance of MAM
Malnutrition is a global crisis (UN SDG2: Zero Hunger) \cite{UN_SDG2}.
% Definition of MAM
Malnutrition refers to deficiencies, excesses, or imbalances in a person’s intake of energy and/or nutrients and can refer to undernutrition, overweight, or obesity.
% Why MAM and not SAM
Moderate Acute Malnutrition (MAM) is the most common, most treatable form of undernutrition.
% What does MAM do
MAM in childhood delays the development of multiple systems in the body from the gut microbiome to neurocognition.
% Treament rates
A small proportion of children with MAM are treated.
% What is treatment
Treatment involves refeeding.
% Problems with refeeding
Current Refeeding methods focus solely on short term weight gain, associate with serious adverse events (such as diarrhoea or Refeeding Syndrome), and are often ineffective at restoring developmental trajectories.
Those that recover their weight often have lasting neurological deficits including cognition and emotional regulation.
% Microbiome connection
MAM is associated with gut microbiome immaturity.
% Microbiome targetted refeeding
Novel prototypical refeeding methods, such as microbiota-directed complementary foods (MDCFs) target the gut-microbiome-brain axis as a mediator to address these deficits.
% Effectiveness of MCDF
By inoculating gnotobiotic pig and mice preclinical models with age-characteristic gut microorganisms, MDCFs increased the levels of biomarkers and mediators of growth, bone formation, neurodevelopment, and immune function toward a state resembling healthy children \cite{gehrig2019effects}.
% Need to improve on these methods
Developing these methods into sustainable refeeding strategies is needed urgently.
% The difficulty in predicting Recovery measurements
% Our study aims
Herein we discuss the results of a study that empoys The Enhanced Ready to Use Supplementary Food (E-RUTF) plus Small Quantity Lipid Based Nutrient Supplement (E-SQLNS) as compared to standard RUTF commonly used to treat childhood Moderate Acute Malnutrition (MAM)

@@

\subsection{What was associated with recovery}
% A background to the data collected
Acknowledging that the refeed type made little impact on the overall recovery rates, an investigation into the factors that were associated with anthropometric recovery was undertaken.
% risk factors for sustained recovery
% are the same factors important for reocvery rates/days to recovery/sustained recovery?
% Why mixed models - maaslin3

@@

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.6]{figures/newrecoveryfeed3.pdf}
	\caption[Refeed methods improve/deteriorate bodily systems differently]{
		Refeed methods improve/deteriorate bodily systems differently.
		A, Radial plot of percentage change in distance from Well-nourished controls before and after refeeding (1yr and 2yr). Positive and negative percentages represent improvement and deterioration, respectively.
		B, Horizontal barplot of features that explain the change significantly (q < 0.2) between refeed methods after mixed modelling for each dataset.
		C, Heatmap to show clusters of dataset changes.
		D, Boxplot to show how Lactobacillus fermentum changes according to the refeed method.
	}
\label{Figure2}
\end{figure}

>>>>

\documentclass{article}
\usepackage{float}
\usepackage[a4paper, top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage[super,sort&compress]{natbib}
\bibliographystyle{unsrtnat}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{booktabs}

@@

\newpage
\section{Main}
% Definition of MAM
Malnutrition refers to deficiencies, excesses, or imbalances in a person’s intake of energy and/or nutrients and can refer to undernutrition, overweight, or obesity.
% Importance of MAM
Malnutrition is a global crisis (UN SDG2: Zero Hunger) \cite{UN_SDG2}.
% Why MAM and not SAM
Moderate Acute Malnutrition (MAM) is the most common, most treatable form of undernutrition.
% What does MAM do
MAM in childhood delays the development of multiple systems in the body from the gut microbiome to neurocognition.
% Treament rates
A small proportion of children with MAM are treated.
% What is treatment
Treatment involves refeeding.
% Problems with refeeding
Current Refeeding methods focus solely on short term weight gain, associate with serious adverse events (such as diarrhoea or Refeeding Syndrome), and are often ineffective at restoring developmental trajectories.
Those that recover their weight often have lasting neurological deficits including cognition and emotional regulation.
% Microbiome connection
MAM is associated with gut microbiome immaturity.
% Microbiome targetted refeeding
Novel prototypical refeeding methods, such as microbiota-directed complementary foods (MDCFs) target the gut-microbiome-brain axis as a mediator to address these deficits.
% Effectiveness of MCDF
By inoculating gnotobiotic pig and mice preclinical models with age-characteristic gut microorganisms, MDCFs increased the levels of biomarkers and mediators of growth, bone formation, neurodevelopment, and immune function toward a state resembling healthy children \cite{gehrig2019effects}.
% Need to improve on these methods
Developing these methods into sustainable refeeding strategies is needed urgently.
% The difficulty in predicting Recovery measurements
% Our study aims
Herein we discuss the results of a study that empoys The Enhanced Ready to Use Supplementary Food (E-RUTF) plus Small Quantity Lipid Based Nutrient Supplement (E-SQLNS) as compared to standard RUTF commonly used to treat childhood Moderate Acute Malnutrition (MAM)

@@

\subsection{What influenced recovery}
% A background to the data collected
Acknowledging that the refeed type made little impact on the overall recovery rates, an investigation into the factors that were associated with anthropometric recovery was undertaken.
% risk factors for sustained recovery
% are the same factors important for reocvery rates/days to recovery/sustained recovery?
% Why mixed models - maaslin3

@@

\subsection{What did recovery influence}
% Maaslin3

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.6]{figures/newrecoveryfeed3.pdf}
	\caption[Refeed methods improve/deteriorate bodily systems differently]{
		Refeed methods improve/deteriorate bodily systems differently.
		A, Radial plot of percentage change in distance from Well-nourished controls before and after refeeding (1yr and 2yr). Positive and negative percentages represent improvement and deterioration, respectively.
		B, Horizontal barplot of features that explain the change significantly (q \textless{} 0.2) between refeed methods after mixed modelling for each dataset.
		C, Heatmap to show clusters of dataset changes.
		D, Boxplot to show how Lactobacillus fermentum changes according to the refeed method.
	}
\label{Figure2}
\end{figure}
